---
title: "Chapter 1 - Introduction to Data"
author: "Chunjie Nan"
output:
    pdf_document:
        extra_dependencies: ["geometry", "multicol", "multirow"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Smoking habits of UK residents**. (1.10, p. 20) A survey was conducted to study the smoking habits of UK residents. Below is a data matrix displaying a portion of the data collected in this survey. Note that "$\pounds$" stands for British Pounds Sterling, "cig" stands for cigarettes, and "N/A" refers to a missing component of the data.

\begin{center}
\scriptsize{
\begin{tabular}{rccccccc}
\hline
	& sex 	 & age 	& marital 	& grossIncome 					     & smoke & amtWeekends	& amtWeekdays \\ 
\hline
1 	& Female & 42 	& Single 	& Under $\pounds$2,600 			     & Yes 	 & 12 cig/day   & 12 cig/day \\ 
2 	& Male	 & 44	& Single 	& $\pounds$10,400 to $\pounds$15,600 & No	 & N/A 			& N/A \\ 
3 	& Male 	 & 53 	& Married   & Above $\pounds$36,400 		     & Yes 	 & 6 cig/day 	& 6 cig/day \\ 
\vdots & \vdots & \vdots & \vdots & \vdots 				             & \vdots & \vdots 	    & \vdots \\ 
1691 & Male  & 40   & Single 	& $\pounds$2,600 to $\pounds$5,200   & Yes 	 & 8 cig/day 	& 8 cig/day \\   
\hline
\end{tabular}
}
\end{center}


(a) What does each row of the data matrix represent?

 Answer: Each row represents  a UK resident who recieved the survey.

(b) How many participants were included in the survey?

 Answer: The matrix shows there are 1691 participants in the survey.

(c) Indicate whether each variable in the study is numerical or categorical. If numerical, identify as continuous or discrete. If categorical, indicate if the variable is ordinal.

 Answer: 
  sex is categorical, age is discrete numerical,  marital is categorical, grossIncome is ordinal categorical, smoke is categorical, amtWeekends is discrete numerical, and amtWeekdays is also discrete numerical.

```{r, eval=FALSE, echo=FALSE}
# NOTE: It is not absolutely necessary to load the data to answer these 
# questions, but is here for further exploration. 
if(!file.exists('11263-Smoking_tcm86-13253.xls')) {
	download.file('https://github.com/jbryer/DATA606Fall2019/blob/master/course_data/11263-Smoking_tcm86-13253.xls?raw=true', 
				  destfile = '11263-Smoking_tcm86-13253.xls')
}
smoking <- readxl::read_excel('11263-Smoking_tcm86-13253.xls')
```





--------------------------------------------------------------------------------

\clearpage

**Cheaters, scope of inference**. (1.14, p. 29) Exercise 1.5 introduces a study where researchers studying the relationship between honesty, age, and self-control conducted an experiment on 160 children between the ages of 5 and 15[^1]. The researchers asked each child to toss a fair coin in private and to record the outcome (white or black) on a paper sheet, and said they would only reward children who report white. Half the students were explicitly told not to cheat and the others were not given any explicit instructions. Differences were observed in the cheating rates in the instruction and no instruction groups, as well as some differences across children's characteristics within each group.

(a) Identify the population of interest and the sample in this study.

 Answer: the population of interest is children and the sample in this study is 160 children between the ages of 5 and 15.

(b) Comment on whether or not the results of the study can be generalized to the population, and
if the findings of the study can be used to establish causal relationships.

 Answer: The causal relationships could be observed from an experiment as the study introduced here,but the researchers told to children reward who toss the white, this possibly cause cheating happens on both who was told not cheating group and  no instruction group as well. In this case, it may difficult to see the facts or results from treatment group base on the control group.Also, researchers should pay attention on the sample size to make sure no bias on the sampling which represents the children between 5 to 15.


[^1]: Alessandro Bucciol and Marco Piovesan. "Luck or cheating? A field experiment on honesty with children". In: Journal of Economic Psychology 32.1 (2011), pp. 73-78. Available at https://papers.ssrn.com/sol3/papers.cfm?abstract_id=1307694



--------------------------------------------------------------------------------

\clearpage

**Reading the paper**. (1.28, p. 31) Below are excerpts from two articles published in the NY Times:

(a) An article titled Risks: Smokers Found More Prone to Dementia states the following:

"Researchers analyzed data from 23,123 health plan members who participated in a voluntary exam and health behavior survey from 1978 to 1985, when they were 50-60 years old. 23 years later, about 25% of the group had dementia, including 1,136 with Alzheimer's disease and 416 with vascular dementia. After adjusting for other factors, the researchers concluded that pack-a- day smokers were 37% more likely than nonsmokers to develop dementia, and the risks went up with increased smoking; 44% for one to two packs a day; and twice the risk for more than two packs."Based on this study, can we conclude that smoking causes dementia later in life? Explain your reasoning.

 Answer: We cannot conclude that smoking causes demential later in life in this case because it is based on observation, not from an experiment, and the study is based on voluntary which can cause a bias on sampling.

(b) Another article titled The School Bully Is Sleepy states the following:

"The University of Michigan study, collected survey data from parents on each child's sleep habits and asked both parents and teachers to assess behavioral concerns. About a third of the students studied were identified by parents or teachers as having problems with disruptive behavior or bullying. The researchers found that children who had behavioral issues and those who were identified as bullies were twice as likely to have shown symptoms of sleep disorders."

A friend of yours who read the article says, "The study shows that sleep disorders lead to bullying in school children." Is this statement justified? If not, how best can you describe the conclusion that can be drawn from this study?

 Answer: I can say that the sleep disorders and bullying in school are associated. However, it's may not a causation problem because we cannot determine that sleep disorder results bulling, and there may have other factors that results or affects bulling in school. 



--------------------------------------------------------------------------------

\clearpage

**Exercise and mental health.** (1.34, p. 35) A researcher is interested in the effects of exercise on mental health and he proposes the following study: Use stratified random sampling to ensure rep- resentative proportions of 18-30, 31-40 and 41-55 year olds from the population. Next, randomly assign half the subjects from each age group to exercise twice a week, and instruct the rest not to exercise. Conduct a mental health exam at the beginning and at the end of the study, and compare the results.

(a) What type of study is this?

 Answer: A blocked random experiment.

(b) What are the treatment and control groups in this study?

 Answer: The treatment group is the group exercise twice a week, and the control group is the group instructed not to exercise.
 
(c) Does this study make use of blocking? If so, what is the blocking variable?

 Answer: Yes, it used blocking, and the blocking variable is age.

(d) Does this study make use of blinding?

 Answer: No, the researchers and the control groups know about the experiment.

(e) Comment on whether or not the results of the study can be used to establish a causal rela- tionship between exercise and mental health, and indicate whether or not the conclusions can be generalized to the population at large.

 Answer: The block random experiment is a good way to conduct the study, but the researchers should have test individuals' mental health before conducting a experiment. If all the samples are at the standard level of mental health, then it is good to generalize the population, but we cannot exclude the situations that some of the sample may had mental illness before recieving the experiment.In this case, we cannot generalize to the population with this experiment.

(f) Suppose you are given the task of determining if this proposed study should get funding. Would you have any reservations about the study proposal?

 Answer: Yes, my reservations about the study is that reseachers should specify what exercise and the amount of exercise the control group should do, and the mental health background of each sample, so that we can see the effects of exercise clearly.



